Status Update regarding \n Assessment of the KLGO coastal waterbird survey monitoring program

### 2020-06-23 analysis/paper/paper.Rmd
minor clean up and edits to overview
Testing - CCRP (AKD)
Making sure I have write access to this repo

# Overview of Progress on this TAR
* Brief overview of activities
  + 2018 Summer - site visit KLGO; discussions w/ Jamie, Elaine Furbish; gather protocol and past report materials, database, etc. Develop assessment strategy and report outline; initial review of past reports and related materials. Recruit 10 hr / wk volunteer undergraduate student intern (Madeleine Ward).
  + 2018 Fall-Winter - bulk of effort ended up focused on working with Madeleine and NPS WASO IT staff to set up computing systems for functional collaborative workflow (e.g., establich NPS github repository and train intern in use of 'version control systems' to automate synchronizing of our files for collaboration, resolving various IT issues, setting up project folder and file structures for long-term use and archiving); initial assessment of program goals and objectives; resolve data access issues; 
  + 2019 Spring - initiate basic summary graphics for clarifying timing and frequency of survey events, variation in duration of survey events (e.g., variation in survey effort at a site); timing and frequency of spp-specific sightings; intial development of phenology graphics. INTERNSHIP ENDED
  + 2019 Summer (minimal time) - prep for August discussion w/ KLGO Resources staff (Jason Taylor, Anne Matsov, Mary Hake); discussed major concern over lack of explicit, specific motivating Objectives (see **Introduction** and **Assessment Results: Phase 1 - Problem Framing and Information Objectives**).
  + 2019 Fall (no progress) - all efforts on hold during Regional Biometrician's detail as NRST Lead
  + 2020 Winter (high level summary of initial concerns added) - in light of the forthcoming transition of the Regional Biometrician to another NPS program, there was a follow up discussion (late Feb) with KLGO Resource Staff (Anne Matsov, Mary Hake) touching on a variety of initial concerns (see **Summary of (Potential) Concerns** below) identified but not yet detailed or assessed quantitatively and thus not yet detailed in the incomplete draft report; conversation ultimately focused on implications of those concerns regarding decisions of (a) potentially reducing or (b) potentially skipping the FY20 field season. Regional Biometrician pointed out that reducing the data collection effort requires further thought and familiarity with existing program's limitations and KLGO's 'hierarchy of objectives' for undertaking this monitoring effort and that, from a statistical perspective, skipping one year in an 15+ year monitoring effort was not problematic in terms of summary analyses of trends and changes.
  + 2020 Feb/Mar (as time permitted) -

@@

## Imporant aspects of the Survey Design 
**Timing**: the start and end dates of the survey vary across years by over a month (see draft Figure 2 below). The timing should be clarified and consistently implemented, as much as possible, to reduce this source of uncertainty and problems in analyzing changes in phenology and other outcomes of interest.  The clarification should be targeted as the priority information goals the Park has identified in the Problem Framing.

@@

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
#git2r::repository("../..")
```

>>>>

# Overview of Progress on this TAR
* Brief overview of activities
  + 2018 Summer - site visit KLGO; discussions w/ Jamie, Elaine Furbish; gather protocol and past report materials, database, etc. Develop assessment strategy and report outline; initial review of past reports and related materials. Recruit 10 hr / wk volunteer undergraduate student intern (Madeleine Ward).
  + 2018 Fall-Winter - bulk of effort ended up focused on working with Madeleine and NPS WASO IT staff to set up computing systems for functional collaborative workflow (e.g., establich NPS github repository and train intern in use of 'version control systems' to automate synchronizing of our files for collaboration, resolving various IT issues, setting up project folder and file structures for long-term use and archiving); initial assessment of program goals and objectives; resolve data access issues; 
  + 2019 Spring - initiate basic summary graphics for clarifying timing and frequency of survey events, variation in duration of survey events (e.g., variation in survey effort at a site); timing and frequency of spp-specific sightings; intial development of phenology graphics. INTERNSHIP ENDED
  + 2019 Summer (minimal time) - prep for August discussion w/ KLGO Resources staff (Jason Taylor, Anne Matsov, Mary Hake); discussed major concern over lack of explicit, specific motivating Objectives (see **Introduction** and **Assessment Results: Phase 1 - Problem Framing and Information Objectives**).
  + 2019 Fall (no progress) - all efforts on hold during Regional Biometrician's detail as NRST Lead
  + 2020 Winter (high level summary of initial concerns added) - in light of the forthcoming transition of the Regional Biometrician to another NPS program, there was a follow up discussion (late Feb) with KLGO Resource Staff (Anne Matsov, Mary Hake) touching on a variety of initial concerns (see **Summary of (Potential) Concerns** below) identified but not yet detailed or assessed quantitatively and thus not yet detailed in the incomplete draft report; conversation ultimately focused on implications of those concerns regarding decisions of (a) potentially reducing or (b) potentially skipping the FY20 field season. Regional Biometrician pointed out that reducing the data collection effort requires further thought and familiarity with existing program's limitations and KLGO's 'hierarchy of objectives' for undertaking this monitoring effort and that, from a statistical perspective, skipping one year in an 15+ year monitoring effort was not problematic in terms of summary analyses of trends and changes.

@@

## Imporant aspects of the Survey Design 
**Timing**: the start and end dates of the survey vary across years by over a month (see draft Figure 2 below). The timing should be clarified and consistently implemented, as much as possible, to reduce this source of uncertainty and problems in analyzing changes in phenology and other outcomes of interest.  The clarification should target the priority information goals the Park has identified in the Problem Framing.

@@

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
#git2r::repository("../..")
```
#
